\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2016-11-17 article.tex
New plots in script

However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy. For these plots, the input data frame should also contain a third quantitative variable column. Each case in this quantitative variable column can be of type numeric, type NA, or type ``". In the first example dataset, columns that could be used for this purpose include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear". At this point, any \pkg{ggenealogy} plot can be used.

@@

@@

In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions. While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{bin}) requires some explanation. The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release). As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels. Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

>>>>

However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy. For these plots, the input data frame should also contain a third quantitative variable column. Each case in this quantitative variable column should be of type numeric. In the first example dataset, columns that could be used for this purpose (once subsetted to filter out values of ``'' and NA) include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear". At this point, any \pkg{ggenealogy} plot can be used.

@@

We would like to note that the \code{plotPath()} method can be created with one or two quantitative variables that the user hard-codes. We illustrate this with the \code{sbGeneal} data frame after we subset it to remove observations that do not have values for the quantitative variable ``yield". Then, we can examine the path between the varieties \code{Clark} and \code{Lawrence} and plot how the quantitative varible ``yield" changes along the parent-child relationships of that path. Likewise, we plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path. The output of these two calls to the \code{plotPath()} function can be viewed in Figure \ref{fig:plotPathQV2}.

\begin{Code}
R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
R> sbFiltIG <- dfToIG(sbFilt)
R> pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
R> plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
R> plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
\end{Code}

\begin{figure}[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotPathQV2}}
    \caption{Left: Plotting a path of interest using a new quantitative variable of interest, ``Yield". Right: Plotting a path of interest using two quantitative variables. We see that the varieties \code{Clark} and \code{Lawrence} have lower yields than the varieties in the middle of the path.}
    \label{fig:plotPathQV2}
\end{figure}

@@

In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions. While the first four explicit parameters have been introduced earlier in this paper, the fifth parameter (\code{bin}) requires some explanation. The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, development year). As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels (unless a user specifies a second quantitative variable of interest, as we will demonstrate later). Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).

@@

We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes. We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created by subsetting the soy bean genealogy data to only include observations that contain a value for the quantitative variable ``yield". Then, we can examine the path across the entire genealogical structure and how the quantitative varible ``yield" changes along the parent-child relationships of that path. Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path. The output of these two calls to the \code{plotPathOnAll()} function can be viewed in Figure \ref{fig:plotPathOnAllQV2}.

\begin{Code}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
\end{Code}

\begin{figure}[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotPathOnAllQV2}}
    \caption{Left: Plotting a path of interest across the genealogical structure using a new quantitative variable of interest, ``Yield". Right: Plotting a path of interest across the genealogical structure using two quantitative variables. We see that an increase in development year might correlate to an increase in yield for the varieties in this dataset, and that our path of interest is composed of generally younger varieties compared to the dataset as a whole.}
    \label{fig:plotPathOnAllQV2}
\end{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

We would like to state that users can still hard code one or two variables in the \code{plotPathOnAll()} function, even with the \code{animate} option set to a value of ``TRUE". The first call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis. When hovering over labels, the label name and yield value is revealed. The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis. When hoering over labels, the label name, yield value, and development year is revealed.

\begin{Code}
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange", animate = TRUE)
R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange", animate = TRUE)
\end{Code}
